---
title: "Big Thicket Ant Analysis"
author: "Marion Donald"
date: "6/3/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
library(knitr)
library(tidyverse)
library(readxl)
library(dplyr)
library(tidyr)
library(vegan)
library(ggplot2)
library(BiodiversityR)
library(extrafont)
library(ggpubr)
library(broom)
library(gridExtra)
library(cowplot)
library(ggpubr)
library(wesanderson)
library(bbmle)
library(lme4)
```


```{r load data, echo = F, warning = F}



data_MH <- read_excel("Meghan Hager Big Thicket Ant Data Updated 07-22-19 (1).xlsx")

data_GZ_all <- read_excel("big_thicket_ants20142018.xlsx", sheet = "Pitfall Data")

native_classification <- read.csv("ant_species_native_classification.csv") 

## GIS dataframe created by Meghan
GIS_df <- read_excel("BIG THICKET GIS DATA 4-6-2016.xlsx", sheet="Sheet1")
```


```{r data clean up, echo = F, warning=F}

data_MH_2 <- data_MH %>% 
  filter(`Collection Method` == "Pit") %>% ## select just the pitfall trap data
  select("Aphenogaster carolinensis":"Trachymyrmex septentrionalis") ## select just the species names

## since MH data doesn't have as many species names as GZ data, we need to add these columns in, in order to merge the two dfs

## get list of species from GZ 
data_GZ_species <- data_GZ_all %>% 
                   filter(Genus != "NA") %>% 
                   unite("species", c("Genus", "Species"), remove = T, sep = " ") %>% 
  select(species) %>% 
  distinct(species) %>% 
  arrange(species) ## finding out that there are some data frame issues -- some of the species names are capitalized while others are not
## this is resulting in duplicates

data_GZ_clean <- data_GZ_all %>% 
  filter(Year == 2014 | ## pull out just the two years of Gabriela's data
           Year == 2015,
         Genus != "NA") %>% 
  mutate(Species = ifelse(Species == "Carolinensis", "carolinensis", ## correct the species names (mispellings and uppercase-to-lowercase)
                          ifelse(Species == "Depilis", "depilis", 
                          ifelse(Species == "Patagonicus", "patagonicus",
                          ifelse(Species == "Castaneus", "castaneus",
                          ifelse(Species == "pennsylanicus" | Species == "Pennsylvanicus" | Species == "pennsylvancius", "pennsylvanicus", 
                          ifelse(Species == "ashmeadii", "ashmeadi", 
                          ifelse(Species == "Rimosus", "rimosus",
                          ifelse(Species == "Opacior", "opacior",
                          ifelse(Species == "Coecus", "coecus", 
                          ifelse(Species == "Americana", "americana", 
                          ifelse(Species == "fasionensis", "faisonensis", 
                          ifelse(Species == "Fulva", "fulva",
                          ifelse(Species == "Harpax", "harpax",
                          ifelse(Species == "Dentata", "dentata",
                          ifelse(Species == "Dentigula", "dentigula",
                          ifelse(Species == "Metallescens", "metallescens",
                          ifelse(Species == "Invicta", "invicta", 
                          ifelse(Species == "Molesta", "molesta",
                          ifelse(Species == "Louisianae", "louisianae", Species))))))))))))))))))),
         Genus = ifelse(Genus == "Pachydondyla", "Pachycondyla", Genus))

data_GZ_clean_sp_check <- data_GZ_clean %>% 
  filter(Abundance != "NA") %>% 
  filter(Genus != "NA") %>% 
  unite("species", c("Genus", "Species"), remove = T, sep = " ") %>% 
  spread("species", "Abundance") %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  summarize_at(vars(`Aphaenogaster carolinensis`:`Trachymyrmex septentrionalis`), sum) %>% 
  gather(species, Abundance)
  
  
data_GZ_clean_sp <- data_GZ_clean_sp_check %>% 
  ungroup() %>% 
  select(species) %>% 
  distinct(species) %>% 
  arrange(species) 



## check species names in MH data **** THERE ARE SPELLING MISTAKES HERE
data_MH_clean <- data_MH %>% 
  rename("Aphaenogaster carolinensis" = 'Aphenogaster carolinensis',
         "Aphaenogaster texana" = "Aphenogaster texana",
         "Hypoponera opaciceps" = "Hyponera opaciceps",
         "Cyphomyrmex rimosus" = "Cyphomyrmex rimosous")

data_MH_sp <- data_MH_clean %>% 
  gather("Aphaenogaster carolinensis":"Trachymyrmex septentrionalis", key = "Species", value = "Abundance") %>% 
  select(Species) %>% 
  distinct(Species) %>% 
  arrange(Species) %>% 
  rename(species = Species)

full_sp_list <- full_join(data_GZ_clean_sp, data_MH_sp, by = "species") ## this is the full species list across the two datasets (but including all data from 2014-2018)
```

```{r species accumulation curve, echo = F, warning =F}

#### data set up for species accumulation curve
##  transform to wide df for species accumulation curve 

data_GZ_wide <- data_GZ_clean %>% 
  filter(Abundance != "NA") %>% 
  select(Year, Site, Station, Month, Genus, Species, Abundance) %>% 
    unite("species", c("Genus", "Species"), remove = T, sep = " ") %>% 
  # group_by(species) %>% 
  #  mutate(grouped_id = row_number()) %>% 
  spread(key = "species", value = "Abundance") %>% 
  mutate_all(~replace(., is.na(.), 0)) %>%  ## make all the NAs under the species names zeros (for this analysis) 
  select(-Site:-Month) ## drop the identifiers and just keep the species names and Year


## find the species that are in MH that we need to add to this df in order to merge them
missing_GZ <- anti_join(data_MH_sp, data_GZ_clean_sp, by = "species") ## 5 species that are in MH that need to be added to GZ

data_GZ_wide2 <- data_GZ_wide %>% 
                mutate("Aphaenogaster texana" = 0,
                       "Brachymyrmex patagonicus" = 0,
                       "Nylanderia terricola" = 0,
                       "Pheidole moerens" = 0,
                       "Prenolepis imparis" = 0)

missing_MH <- anti_join(data_GZ_clean_sp, data_MH_sp, by = "species") ## 14 spp that are in GZ that need to be added to MH

data_MH_2a <- data_MH_clean %>% 
  mutate("Aphaenogaster rudis-fulva-texana complex" = 0,
         "Aphaenogaster treatae" = 0,
         "Brachymyrmex depilis" = 0,
         "Crematogaster ashmeadi" = 0,
         "Crematogaster lineolata" = 0,
         "Crematogaster minutissima" = 0,
         "Crematogaster sp" = 0,
         "Hypoponera opacior" = 0,
         "Labidus coecus" = 0,
         "Pheidole flavens complex" = 0,
         "Pheidole sp" = 0, 
         "Pseudomyrmex pallidus" = 0, 
         "Solenopsis nickersoni" = 0, 
         "Tapinoma sessile" = 0) %>% 
  select("Aphaenogaster carolinensis":"Tapinoma sessile",
         -Site_Code, -`Inside or Outside (Inside=1)`) %>% 
  mutate(Year = 2015)

data_spacc_all <- data_MH_2a %>% 
  bind_rows(data_GZ_wide2) %>% 
  mutate(`Pheidole flavens complex` = `Pheidole flavens complex` + `Pheidole moerens`) %>% ## combining P. moerens since it can't be reliably identified from P. flavens complex
  select(-`Pheidole moerens`) ## drop P. moerens now that it's been combined with P. flavens


data_spacc_year <- data_spacc_all %>% 
                  select(Year)

data_spacc <- data_spacc_all %>% 
              select(-Year) # drop year, just keep the species by site matrix
curve <- specaccum(data_spacc, method = "random")

pitfalls <- as.data.frame(curve$sites)
richness <- as.data.frame(curve$richness)
sd <- as.data.frame(curve$sd)

spacc_df <- pitfalls %>% 
  bind_cols(richness, sd) %>% 
  dplyr::rename(sites = `curve$sites`, 
         richness = `curve$richness`,
         sd = `curve$sd`)

sp_accumulation_curve <- ggplot(spacc_df, aes(sites, richness))+
  geom_point(alpha = .5)+
  geom_ribbon(aes(x = sites, ymin = richness-sd, ymax = richness+sd), alpha = 0.2)+
  theme_classic()+
  labs(x = "Number of pitfall traps",
       y = "Species richness")+
  xlim(c(0, 410))

## Chao estimated (36 species observed, ~42 +/- 6 species predicted by Chao, 42 predicted by jack1 +/- 2)
richness_est_df <- vegan::specpool(data_spacc)
```
## Species accumulation curve richness estimates
```{r species accumulation figure, echo = F,fig.width = 3.5, fig.height=2.5, fig.align = "center"}
sp_accumulation_curve
```

Species accumulation curve from 381 pitfall traps, collected by Gabriela and Meghan. Here, we used the function 'specaccum' from the vegan package in R, which estimates the number of species for a given number of sampled sites. Richness estimates are shown as points with +/- the standard deviation as the shaded ribbon.
We then extrapolated the number of unobserved species, using the function 'specpool' from the vegan package in R. We compared this estimate of unobserved species with the number of observed species to determine the efficacy of our sampling. 
```{r richness estimates, echo =F}
kable(richness_est_df)
```

```{r rank abundance, echo = F, warning = F}

## rank abundance data frame
data_rank_abundance <- data_spacc_all %>% 
  summarize_at(vars(`Aphaenogaster carolinensis`:`Tapinoma sessile`), sum) %>% 
  gather(species, abundance) %>% 
  arrange(-abundance) %>% 
  rownames_to_column(var = "rank") %>% 
  mutate(rank = as.numeric(rank))

## native classification df to merge (native = 1, invasive = 0)
native_info <- native_classification %>% 
  rename(species = "Species") %>% 
  mutate(species = as.character(species))


rank_abu_df2 <- data_rank_abundance  %>% 
  full_join(native_info, by = "species") %>% 
  mutate(Native = as.character(Native))
         
## Pheidole flavens complex as Native for rank abundance
rank_abu_P.flav.native <- rank_abu_df2 %>% 
  mutate(Native = ifelse(species == "Pheidole flavens complex", 1, Native),
         Classification = ifelse(Native == 1, "native", "non-native")) 



## Rank abundance curve with Pheidole flavens complex as native (1)
rank_abu_fig <- ggplot(rank_abu_P.flav.native, aes(x=rank, y = abundance, label=species))+
  geom_point(aes(color = Classification, shape = Classification), size = 4)+
  geom_line()+
  theme_classic()+
  labs(x = "Species rank",
       y = "Abundance")+
  geom_text(aes(label=ifelse(abundance>100,as.character(species),'')),hjust=-0.08,vjust=0, check_overlap = T)+
  scale_color_manual(values = c("#F1BB7B","#5B1A18","#055864","#04C0DD", "gray49"))+
  scale_shape_manual(values = c(16, 16))+
  theme(legend.position = "hide",
        legend.key = element_blank(),
        legend.background=element_blank())

## Pheidole flavens complex as Native for rank abundance
rank_abu_P.flav.non.native <- rank_abu_df2 %>% 
  mutate(Native = ifelse(species == "Pheidole flavens complex", 0, Native),
         Classification = ifelse(Native == 1, "native", "non-native")) 



## Rank abundance curve with Pheidole flavens complex as non-native (NN) (0)
rank_abu_NN_fig <- ggplot(rank_abu_P.flav.non.native, aes(x=rank, y = abundance, label=species))+
  geom_point(aes(color = Classification, shape = Classification), size = 4)+
  geom_line()+
  theme_classic()+
  labs(x = "Species rank",
       y = "Abundance")+
  geom_text(aes(label=ifelse(abundance>100,as.character(species),'')),hjust=-0.08,vjust=0, check_overlap = T)+
  scale_color_manual(values = c("#F1BB7B","#5B1A18", "#055864","#04C0DD", "gray49"))+
  scale_shape_manual(values = c(16, 16))+
  theme(legend.position = "hide",
        legend.key = element_blank(),
        legend.background=element_blank())

```


## Rank abundance curves

```{r rank abu figs, echo = F, fig.width=6, fig.height=3, warning=F}
legend_native_class <- get_legend(rank_abu_fig + theme(legend.position="bottom", legend.key = element_blank(), legend.background=element_blank()))

rank_abu_figs <- plot_grid(rank_abu_fig, rank_abu_NN_fig,  labels = "auto")

plot_grid( rank_abu_figs, legend_native_class, ncol = 1, rel_heights = c(1, .04))
```

To visualize the evenness and relative abundance of the species, we created rank abundance curves. In these two figures, *Pheidole flavens* complex is classified as (a) a native species and (b) a non-native species. (*Scott, I think it makes sense to only show one of these. So depending on how P. flavens is classified, we can use the corresponding figure.*)

```{r alpha diversity, echo = F, warning = F, include=F, warning = F}

## diversity measures -- no difference in alpha diversity across the years
shannon_div <- diversity(data_spacc, "shannon")

shannon_div_df <- as.data.frame(shannon_div) %>% 
                  bind_cols(data_spacc_year) %>% 
  mutate(Year = as.factor(Year))

ggplot(shannon_div_df, aes(x = Year, y = shannon_div))+
  geom_boxplot()+
  geom_jitter(width = 0.05, height = 0)+
  theme_classic()+
  labs(x = "Sampling year",
       y = "Shannon diversity index")+
  stat_compare_means(method = "t.test")

## simpson div - no difference in alpha diversity across the years
simpson_div <- diversity(data_spacc, "simpson")

simpson_div_df <- as.data.frame(simpson_div) %>% 
  bind_cols(data_spacc_year) %>% 
  mutate(Year = as.factor(Year))

ggplot(simpson_div_df, aes(x = Year, y = simpson_div))+
  geom_boxplot()+
  geom_jitter(width = 0.05, height = 0)+
  theme_classic()+
  labs(x = "Sampling year",
       y = "Simpson diversity index")+
  stat_compare_means(method = "t.test")

## get df with info on sample ID (from row number) and year to get proportion of non-native to native ants (and also do it as occurrence for sp)
data_spacc_all_tidy <- data_spacc_all %>% 
  rownames_to_column("ID") %>% 
  gather(key = "species", value = "abundance", c(-Year, -ID)) %>% 
  full_join(native_info)

## Pheidole flavens complex is classed as "2" -- convert to native (1) or invasive (0) and run the analyses for proportion of invasives
data_spacc_all_tidy_PF_native <- data_spacc_all_tidy %>% 
  mutate(Native = ifelse(species == "Pheidole flavens complex", 1, Native))

### do the same as above but this time by species occurrence (presence/absence) rather than abundance (Pheidole flavens complex as native)
data_spacc_all_tidy_occ <- data_spacc_all_tidy_PF_native %>% 
  filter(abundance != 0) %>%  # sp has to be present to contribute
 filter(Native == 1) %>% 
  group_by(ID, Year) %>% 
  summarize(num_native_sp = n())

data_spacc_all_tidy_non_native <- data_spacc_all_tidy %>% 
  filter(abundance != 0) %>%  # sp has to be present to contribute
  filter(Native == 0) %>% 
  group_by(ID, Year) %>% 
  summarize(num_nonnative_sp = n())

occurrence_proportion_df <- data_spacc_all_tidy_occ %>% 
  full_join(data_spacc_all_tidy_non_native) %>% 
  mutate_at(vars(-group_cols()),~replace(., is.na(.), 0)) %>% 
  mutate(total_sp = num_native_sp + num_nonnative_sp,
         prop_nonnative_sp = num_nonnative_sp/total_sp) %>% 
  ungroup() %>% 
  mutate(ID = as.numeric(ID)) %>% 
  arrange(ID) %>% 
  bind_cols(shannon_div_df, 
            simpson_div_df)


### for shannon div
#  linear model
lm_fit_shan_occ<- lm(shannon_div ~ prop_nonnative_sp, data=occurrence_proportion_df)
#summary(lm_fit_shan_occ)

#plot(lm_fit_shan_occ)
#tidy(lm_fit_shan_occ)

#glance(lm_fit_shan_occ)

## residuals plot looks more "random" -- but still not a great fit
ggplot(occurrence_proportion_df, aes(prop_nonnative_sp, shannon_div))+
  geom_jitter()+
  geom_line(data = broom::augment(lm_fit_shan_occ), aes(x = prop_nonnative_sp, y = .fitted))+
  #facet_grid(~Year)+
  theme_classic()+
  labs(x = "Proportion of non-native ant species",
       y = "Shannon diversity index")

### for simpson div
#  linear model
lm_fit_simp_occ<- lm(simpson_div ~ prop_nonnative_sp, data=occurrence_proportion_df)
summary(lm_fit_simp_occ)

#plot(lm_fit_simp_occ)
#tidy(lm_fit_simp_occ)

#glance(lm_fit_simp_occ)


## residuals plot looks more "random" -- but not a great fit
prop_simpson_lm <- ggplot(occurrence_proportion_df, aes(prop_nonnative_sp, simpson_div))+
  geom_jitter()+
 #geom_smooth(se = F)+
  geom_line(data = broom::augment(lm_fit_simp_occ), aes(x = prop_nonnative_sp, y = .fitted))+
  #facet_grid(~Year)+
  theme_classic()+
  labs(x = "Proportion of non-native ant species",
       y = "Simpson diversity index")

prop_simpson_lm_loess <- ggplot(occurrence_proportion_df, aes(prop_nonnative_sp, simpson_div))+
  geom_jitter()+
  geom_smooth(se = F)+
  geom_line(data = broom::augment(lm_fit_simp_occ), aes(x = prop_nonnative_sp, y = .fitted))+
  #facet_grid(~Year)+
  theme_classic()+
  labs(x = "Proportion of non-native ant species",
       y = "Simpson diversity index")

occurrence_proportion_df_cat <- occurrence_proportion_df %>% 
  filter(total_sp != 0) %>% 
  mutate(category = ifelse(prop_nonnative_sp < .33, "Low", 
                           ifelse(prop_nonnative_sp > .66, "High", "Medium")))

occurrence_proportion_df_cat$category <- factor(occurrence_proportion_df_cat$category, levels = c("Low", "Medium", "High"))

my_comparisons <- list(   c("High", "Medium"),  c("Low", "Medium"), c("Low", "High") )

## get species richness (total # sp per trap) to use as the species richness diversity measure by abundance later
sp_richness<- occurrence_proportion_df_cat %>% 
  select(total_sp, ID)

occurrence_proportion_df_cat_tidy <- occurrence_proportion_df_cat %>% 
  gather(diversity_type, diversity_measure, -c(ID, Year, num_native_sp, num_nonnative_sp,
                                               prop_nonnative_sp, Year1, Year2, category)) %>% 
  mutate(diversity_type = case_when(diversity_type == "shannon_div" ~ "Shannon diversity index",
                                   diversity_type == "simpson_div" ~ "Simpson diversity index",
                                   diversity_type == "total_sp" ~ "Species richness"))

occurrence_prop_sp_fig <- ggplot(occurrence_proportion_df_cat_tidy, aes(category, diversity_measure, fill = category))+
  geom_boxplot(position =position_dodge())+
  geom_point(position = position_jitterdodge(0.25), alpha = .75)+
  stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..))+
  scale_fill_manual(values = c("#F1BB7B", "#FD6467", "#5B1A18", "#D67236"))+
  scale_color_manual(values = c("#F1BB7B", "#FD6467", "#5B1A18", "#D67236"))+
  labs(y = "Alpha diversity",
       x = "Proportion of non-native species (by occurrence with P.flavens as native)")+
  theme_classic()+
  theme(#text = element_text(family = "Times New Roman", size = 14),
        legend.position = "bottom",
        legend.title = element_blank())+
  facet_wrap(diversity_type~., scales="free_y")




##### P. flavens as native -- proportion non-native analysis
data_spacc_all_tidy_abu <- data_spacc_all_tidy_PF_native %>% 
  group_by(ID, Year, Native) %>% 
  summarize(abundance = sum(abundance))

total_abu <- data_spacc_all_tidy_abu %>% 
  ungroup() %>% 
  group_by(ID, Year) %>% 
  summarize(total_abundance = sum(abundance))

non_native_abu <- data_spacc_all_tidy_abu %>% 
  filter(Native == 0) %>% 
  rename(non_native_abu = abundance) %>% 
  select(-Native)

non_native_proportion_df <- total_abu %>% 
  left_join(non_native_abu, by = c("ID", "Year")) %>% 
  filter(total_abundance != "NA") %>% 
  mutate(non_native_abu_prop = non_native_abu/total_abundance) %>% 
  ungroup() %>% 
  mutate(ID = as.numeric(ID)) %>% 
  arrange(ID) %>% 
  bind_cols(shannon_div_df,
            simpson_div_df)

non_native_proportion_df_cat <- non_native_proportion_df %>% 
  filter(total_abundance != 0) %>% 
  mutate(category = ifelse(non_native_abu_prop < .33, "Low", 
                           ifelse(non_native_abu_prop > .66, "High", "Medium"))) %>% 
  left_join(sp_richness, by = "ID")

non_native_proportion_df_cat$category <- factor(non_native_proportion_df_cat$category, levels = c("Low", "Medium", "High"))

my_comparisons <- list(   c("High", "Medium"),  c("Low", "Medium"), c("Low", "High") )



non_native_proportion_df_cat_tidy <- non_native_proportion_df_cat %>% 
  select(ID, Year, non_native_abu_prop, shannon_div, simpson_div, total_sp, category) %>% 
  gather(diversity_type, diversity_measure, -c(ID, Year, non_native_abu_prop, category)) %>% 
  mutate(diversity_type = case_when(diversity_type == "shannon_div" ~ "Shannon diversity index",
                                    diversity_type == "simpson_div" ~ "Simpson diversity index",
                                    diversity_type == "total_sp" ~ "Species richness"))

abundance_prop_alpha_fig <- ggplot(non_native_proportion_df_cat_tidy, aes(category, diversity_measure, fill = category))+
  geom_boxplot(position =position_dodge())+
  geom_point(position = position_jitterdodge(0.25), alpha = .75)+
  stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..))+
  scale_fill_manual(values = wes_palette("GrandBudapest1"))+
  scale_color_manual(values = wes_palette("GrandBudapest1"))+
  labs(y = "Alpha diversity",
       x = "Proportion of non-native species (by abundance P.flavens as native)")+
  theme_classic()+
  theme(#text = element_text(family = "Times New Roman", size = 14),
        legend.position = "bottom",
        legend.title = element_blank())+
  facet_wrap(diversity_type~., scales="free_y")

#  linear model
lm_fit <- lm(shannon_div ~ non_native_abu_prop, data=non_native_proportion_df)
summary(lm_fit)
plot(lm_fit) ## residuals are not "random" -- they form an inverted U, linear model is not a great fit for these data
tidy(lm_fit)

glance(lm_fit)

## looks like there's a hump-shaped relationship between alpha diversity and proportion of non-native ants,
## not sure that the linear model is the best fit for these data
ggplot(non_native_proportion_df, aes(non_native_abu_prop, shannon_div))+
  geom_point()+
  #geom_line(data = broom::augment(lm_fit), aes(x = non_native_abu_prop, y = .fitted))+
  #facet_grid(~Year)+
  theme_classic()+
    labs(x = "Proportion of non-native ant abundance",
         y = "Shannon diversity index")

### for simpson div
#  linear model
lm_fit_simp<- lm(simpson_div ~ non_native_abu_prop, data=non_native_proportion_df)
#summary(lm_fit_simp)

#plot(lm_fit_simp)
#tidy(lm_fit_simp)

#glance(lm_fit_simp)

## looks like there's a hump-shaped relationship between alpha diversity and proportion of non-native ants,
## not sure that the linear model is the best fit for these data
ggplot(non_native_proportion_df, aes(non_native_abu_prop, simpson_div))+
  geom_point()+
  geom_line(data = broom::augment(lm_fit_simp), aes(x = non_native_abu_prop, y = .fitted))+
  #facet_grid(~Year)+
  theme_classic()+
  labs(x = "Proportion of non-native ant abundance",
       y = "Shannon diversity index")
```

## Alpha diversity -- local species richness
The original figure compared the proportion of non-native ant species out of total species with Shannon and Simpson diversity indices, and used a linear model to test for a correlation between the two. I think this might not be the best way to visualize these data or to test for a trend between alpha diversity and proportion of non-native ants, due to the non-linear spread of the data. Looking at this figure (a), richness appears to peak at an intermediate level of non-native proportion of total ant species. Alternatively, we could use the default loess model (b), which shows the peak at an intermediate proportion of non-native to total ant species.  

```{r alpha div example switch, echo = F, warning = F, fig.height=2.5}
plot_grid(prop_simpson_lm, prop_simpson_lm_loess, labels = "auto")
```

However, I think a better way to visualize these data is to bin the proportion of non-native ant species (or abundances) out of total species (or abundance) and compare alpha diversity across these categories using Wilcoxon tests. 

Basing the proportion of non-native to total species by occurrence and abundance gives qualitatively the same result -- alpha diversity peaks at an intermediate level of non-native species, and is lowest when the proportion of non-native species is high. This remains true whether or not *P. flavens* complex is classified as native or non-native. 

```{r alpha div P. flav native, echo = F, warning = F, fig.height=8, warning = F}
## P. flavens complex as native
plot_grid(occurrence_prop_sp_fig,abundance_prop_alpha_fig, nrow=2, labels = "auto")
```

These are the proportion of non-native species (with *P. flavens* complex as native) by (a) occurrence and (b) abundance, binned them by low (0-.33) medium (.34-.66) and high (.67-1.0), and compared these categories with three difference measures of local diversity (Shannon, Simpson, and species richness). Qualitatively, the trends are similar between the two sets of figures.
```{r alpha div P.flav non-native, echo = F, warning = F, include = F}

##### RUN THE ABOVE WITH P. flavens as non-native
data_spacc_all_tidy_PF_non_native <- data_spacc_all_tidy %>% 
  mutate(Native = ifelse(species == "Pheidole flavens complex", 0, Native))


data_spacc_all_tidy_abu_NN <- data_spacc_all_tidy_PF_non_native %>% 
  group_by(ID, Year, Native) %>% 
  filter(abundance != "NA") %>% 
  summarize(abundance = sum(abundance))

total_abu_NN <- data_spacc_all_tidy_abu_NN %>% 
  ungroup() %>% 
  group_by(ID, Year) %>% 
  summarize(total_abundance = sum(abundance))

non_native_abu_NN <- data_spacc_all_tidy_abu_NN %>% 
  filter(Native == 0) %>% 
  rename(non_native_abu = abundance) %>% 
  select(-Native)

non_native_proportion_df_NN <- total_abu_NN %>% 
  left_join(non_native_abu_NN) %>% 
  mutate(non_native_abu_prop = non_native_abu/total_abundance) %>% 
  ungroup() %>% 
  mutate(ID = as.numeric(ID)) %>% 
  arrange(ID) %>% 
  bind_cols(shannon_div_df,
            simpson_div_df)


non_native_proportion_df_NN_cat <- non_native_proportion_df_NN %>% 
  filter(total_abundance != 0) %>% 
  mutate(category = ifelse(non_native_abu_prop < .33, "Low", 
                           ifelse(non_native_abu_prop > .66, "High", "Medium"))) %>% 
  left_join(sp_richness, by = "ID")

non_native_proportion_df_NN_cat$category <- factor(non_native_proportion_df_NN_cat$category, levels = c("Low", "Medium", "High"))

my_comparisons <- list(   c("High", "Medium"),  c("Low", "Medium"), c("Low", "High") )



non_native_proportion_df_NN_cat_tidy <- non_native_proportion_df_NN_cat %>% 
  select(ID, Year, non_native_abu_prop, shannon_div, simpson_div, total_sp, category) %>% 
  gather(diversity_type, diversity_measure, -c(ID, Year, non_native_abu_prop, category)) %>% 
  mutate(diversity_type = case_when(diversity_type == "shannon_div" ~ "Shannon diversity index",
                                    diversity_type == "simpson_div" ~ "Simpson diversity index",
                                    diversity_type == "total_sp" ~ "Species richness"))

abundance_prop_alpha_fig_NN <- ggplot(non_native_proportion_df_NN_cat_tidy, aes(category, diversity_measure, fill = category))+
  geom_boxplot(position =position_dodge())+
  geom_point(position = position_jitterdodge(0.25), alpha = .75)+
  stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..))+
  scale_fill_manual(values = wes_palette("GrandBudapest1"))+
  scale_color_manual(values = wes_palette("GrandBudapest1"))+
  labs(y = "Alpha diversity",
       x = "Proportion of non-native species (by abundance with P.flavens as non-native)")+
  theme_classic()+
  theme(#text = element_text(family = "Times New Roman", size = 14),
        legend.position = "bottom",
        legend.title = element_blank())+
  facet_wrap(diversity_type~., scales="free_y")


#  linear model
lm_fit <- lm(shannon_div ~ non_native_abu_prop, data=non_native_proportion_df_NN)
summary(lm_fit)
#plot(lm_fit) ## residuals are not "random" -- they form an inverted U, linear model is not a great fit for these data
#tidy(lm_fit)

#glance(lm_fit)

## looks like there's a hump-shaped relationship between alpha diversity and proportion of non-native ants,
## not sure that the linear model is the best fit for these data
ggplot(non_native_proportion_df, aes(non_native_abu_prop, shannon_div), color = "red")+
  geom_jitter(data = non_native_proportion_df, color = "black")+
  geom_point(data= non_native_proportion_df_NN, aes(color = "red"))+
  geom_line(data = broom::augment(lm_fit), aes(x = non_native_abu_prop, y = .fitted))+
  #facet_grid(~Year)+
  theme_classic()+
  labs(x = "Proportion of non-native ant abundance",
       y = "Shannon diversity index")

### for simpson div
#  linear model
lm_fit_simp<- lm(simpson_div ~ non_native_abu_prop, data=non_native_proportion_df)
summary(lm_fit_simp)

plot(lm_fit_simp)
tidy(lm_fit_simp)

glance(lm_fit_simp)

## looks like there's a hump-shaped relationship between alpha diversity and proportion of non-native ants,
## not sure that the linear model is the best fit for these data
ggplot(non_native_proportion_df, aes(non_native_abu_prop, simpson_div))+
  geom_point()+
  geom_line(data = broom::augment(lm_fit), aes(x = non_native_abu_prop, y = .fitted))+
  #facet_grid(~Year)+
  theme_classic()+
  labs(x = "Proportion of non-native ant abundance",
       y = "Shannon diversity index")

### do the same as above but this time by species occurrence (presence/absence) rather than abundance (Pheidole flavens complex as native)
data_spacc_all_tidy_occ_NN <- data_spacc_all_tidy_PF_non_native %>% 
  filter(abundance != 0) %>%  # sp has to be present to contribute
  filter(Native == 1) %>% 
  group_by(ID, Year) %>% 
  summarize(num_native_sp = n())

data_spacc_all_tidy_non_native <- data_spacc_all_tidy_PF_non_native %>% 
  filter(abundance != 0) %>%  # sp has to be present to contribute
  filter(Native == 0) %>% 
  group_by(ID, Year) %>% 
  summarize(num_nonnative_sp = n())

occurrence_proportion_df_NN <- data_spacc_all_tidy_occ_NN %>% 
  full_join(data_spacc_all_tidy_non_native) %>% 
  mutate_at(vars(-group_cols()),~replace(., is.na(.), 0)) %>% 
  mutate(total_sp = num_native_sp + num_nonnative_sp,
         prop_nonnative_sp = num_nonnative_sp/total_sp) %>% 
  ungroup() %>% 
  mutate(ID = as.numeric(ID)) %>% 
  arrange(ID) %>% 
  bind_cols(shannon_div_df, 
            simpson_div_df)

occurrence_proportion_df_cat_NN <- occurrence_proportion_df_NN %>% 
  filter(total_sp != 0) %>% 
  mutate(category = ifelse(prop_nonnative_sp < .33, "Low", 
                           ifelse(prop_nonnative_sp > .66, "High", "Medium")))

occurrence_proportion_df_cat_NN$category <- factor(occurrence_proportion_df_cat_NN$category, levels = c("Low", "Medium", "High"))

my_comparisons <- list(   c("High", "Medium"),  c("Low", "Medium"), c("Low", "High") )


occurrence_proportion_df_cat_tidy_NN <- occurrence_proportion_df_cat_NN %>% 
  gather(diversity_type, diversity_measure, -c(ID, Year, num_native_sp, num_nonnative_sp,
                                               prop_nonnative_sp, Year1, Year2, category)) %>% 
  mutate(diversity_type = case_when(diversity_type == "shannon_div" ~ "Shannon diversity index",
                                    diversity_type == "simpson_div" ~ "Simpson diversity index",
                                    diversity_type == "total_sp" ~ "Species richness"))

occurrence_prop_sp_fig_NN <- ggplot(occurrence_proportion_df_cat_tidy_NN, aes(category, diversity_measure, fill = category))+
  geom_boxplot(position =position_dodge())+
  geom_point(position = position_jitterdodge(0.25), alpha = .75)+
  stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..))+
  scale_fill_manual(values = wes_palette("GrandBudapest1"))+
  scale_color_manual(values = wes_palette("GrandBudapest1"))+
  labs(y = "Alpha diversity",
       x = "Proportion of non-native species (by occurrence with P.flavens as non-native)")+
  theme_classic()+
  theme(#text = element_text(family = "Times New Roman", size = 14),
        legend.position = "bottom",
        legend.title = element_blank())+
  facet_wrap(diversity_type~., scales="free_y")
```

```{r alpha div P. flavens complex as non-native, echo = F, warning=F, fig.height = 8}
plot_grid(occurrence_prop_sp_fig_NN, abundance_prop_alpha_fig_NN, nrow =2, labels = "auto")
```

These are the proportion of non-native species (with *P. flavens* complex as non-native) by (a) occurrence and (b) abundance, binned them by low (0-.33) medium (.34-.66) and high (.67-1.0), and compared these categories with three difference measures of local diversity (Shannon, Simpson, and species richness). Qualitatively, the trends are similar between the two sets of figures.


## NMDS

```{r nmds set up, echo=FALSE, include = F}

## Next step is to pull just Gabriela's data and look at year to year trends + seasonality in species composition
## using ordination (nMDS or PCoA and PERMANOVA + permdist)
data_NMDS <- data_GZ_wide %>% 
  select(-Year)




# Then calculate multivariate distances between samples using vegdist() with one of various distance metrics:
dist_jac <- vegdist(data_NMDS, method = "jaccard", binary = TRUE) # Binary jaccard distance (presence/absence)
dist_BC <- vegdist(data_NMDS, method = "bray")#, binary = TRUE) # Bray-Curtis distance

# Then generate an MDS object using metaMDS():
#mds <- metaMDS(dist_BC) ## stress is too low with the full community -- too many super rare species
#str(mds)



## NMDS isn't converging with all of the individual pitfalls -- likely because many of them have few but differenct ant spp
## try this again but now grouped by site within year and season

## drop super rare species (present in > 1% of of total abundance) -- this results in 11 species

data_new <- data_GZ_clean %>% 
  select(Year, Month, Genus, Species, Abundance) %>% 
  filter(Year == 2014 |
           Year == 2015) %>% 
  rownames_to_column() %>% 
  unite(species, c("Genus","Species"), sep =  " ", remove = F) %>% 
  spread(key = "species", value = "Abundance") %>% 
  mutate_all(~replace(., is.na(.), 0))%>%  ## make all the NAs under the species names zeros (for this analysis) 
  group_by(Year, Month) %>% 
  summarize_at(vars(`Aphaenogaster carolinensis`:`Trachymyrmex septentrionalis`), sum) %>% 
  gather(key = "species", value = "Abundance", c(-Year,-Month)) %>% 
  ungroup() %>% 
  group_by(species) %>% 
  summarize(Abundance = sum(Abundance)) %>% 
  mutate(percent_total = Abundance/1754) %>% 
  filter(percent_total > 0.01)

sp_list <- as.list(data_new$species)

total_abu <- data_new %>% 
  ungroup() %>% 
  summarize(total_abundance = sum(Abundance))

data_GZ_clean_subset <- data_GZ_clean  %>% 
  unite(species, c("Genus", "Species"), sep =" ", remove = F) %>% 
filter(species %in% sp_list) %>% 
spread("species", "Abundance") %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  group_by(Year, Month, Site) %>% ## summarize by site since we have multiple pitfall traps per site
  ## and the individual pitfall level resolution is too fine-scale
  summarize_at(vars(`Aphaenogaster carolinensis`:`Solenopsis molesta`), sum)


GZ_wide_NMDS_new <- data_GZ_clean_subset %>% 
  ungroup() %>% 
  select(-c(Year:Site))



# Then calculate multivariate distances between samples using vegdist() with one of various distance metrics:
dist_jac <- vegdist(GZ_wide_NMDS_new, method = "jaccard", binary = TRUE) # Binary jaccard distance (presence/absence)
dist_BC <- vegdist(GZ_wide_NMDS_new, method = "bray")#, binary = TRUE) # Bray-Curtis distance

# Then generate an MDS object using metaMDS(): ### hooray!! it converges and stress is 0.173
mds <- metaMDS(dist_BC, trymax = 300)

# Plot the mds using basic graphics
plot(mds)

# Make this into a ggplot:
# Extract the coordinates:
points <- as.data.frame(mds$points)
points

# Add metadata to the points:
data_GZ_clean_subset_2 <- data_GZ_clean_subset %>% 
  rownames_to_column() %>% 
  mutate(Season = ifelse(Month == 5, "Spring", "Fall")) %>% 
  unite(Season_yr, c(Season, Year), remove =F)

## get the proportion of non-native/total spp for this subset of data
subset_prop_NN <- data_GZ_clean_subset_2 %>% 
  gather("Species", "abundance", -c(rowname:Site, Season)) %>% 
  group_by(Season_yr, Site) %>% 
  left_join(native_classification) %>%
  filter(abundance >0) %>% 
  summarize(species_total = length(abundance),
            native_total = sum(Native)) %>% 
  mutate(non_native_total = species_total - native_total,
         prop_non_native = non_native_total/species_total,
  category = ifelse(prop_non_native < .33, "Low", 
                    ifelse(prop_non_native > .66, "High", "Medium")))


data_GZ_clean_subset_2 <- data_GZ_clean_subset_2 %>% 
  left_join(subset_prop_NN, by = c("Season_yr", "Site"))

points <- points %>% mutate(Sample = rownames(points),
                            Year = data_GZ_clean_subset_2$Year[match(rownames(points), data_GZ_clean_subset_2$rowname)],
                            Month = data_GZ_clean_subset_2$Month[match(rownames(points), data_GZ_clean_subset_2$rowname)],
                            Season = data_GZ_clean_subset_2$Season[match(rownames(points), data_GZ_clean_subset_2$rowname)],
                            Season_yr = data_GZ_clean_subset_2$Season_yr[match(rownames(points), data_GZ_clean_subset_2$rowname)],
                            category = data_GZ_clean_subset_2$category[match(rownames(points), data_GZ_clean_subset_2$rowname)],
                            prop_non_native = data_GZ_clean_subset_2$prop_non_native[match(rownames(points), data_GZ_clean_subset_2$rowname)]) 

points <- points %>% 
  mutate(Year = as.factor(Year))

```
```{r nmds figs, echo = F, warning=F}

nmds_season <- ggplot(data = points, aes(x = MDS1, y = MDS2))+
  geom_point(aes(color = Season))+
  stat_ellipse(aes(color = (Season)))+
  scale_color_manual(values = c("darkgray", "black"))+
  labs(x = "NMDS1",
       y = "NMDS2")+
  theme_classic()+
  theme(legend.position = "bottom")

nmds_year<-ggplot(data = points, aes(x = MDS1, y = MDS2))+
  geom_point(aes(color = Year))+
  stat_ellipse(aes(color = Year))+
  scale_color_manual(values = c("darkgray", "black"))+
  labs(x = "NMDS1",
       y = "NMDS2")+
  theme_classic()+
  theme(legend.position = "bottom")

plot_grid(nmds_season, nmds_year, ncol=2, label ="auto")

```

Visualization of the ant communities (using a bray-curtis dissimilarity matrix). Each point represents the pooled community across pitfall traps within each site. Colors represent (a) seasons and (b) years. 
We can use PERMANOVA analyses to test whether ant community composition varied between seasons and years. PERMANOVA tests for differences in the locations of multivariate groups. We used the adonis function of vegan. As we had repeated community measures within site, we constrained permutations to within sites (using the "strata" argument). We included season and year as fixed effects in the models. These analyses confirmed that season had a significant effect on community composition and explained approximately 5% of the variation ($R^2$ = 0.055)
```{r permanova test, echo = F, warning=F}
perm <- adonis(dist_BC ~ Season*Year, data = data_GZ_clean_subset_2, strata = data_GZ_clean_subset_2$Site)

perm
```

```{r permdist homogeneity, echo = F, warning = F, include = F}

## Composition in ant community affected by seasonality but not by year

## check for homogeneity in dispersion - if there is, it may invalidate the PERMANOVA sp. composition results

group <- data_GZ_clean_subset_2$Month
## Calculate multivariate dispersions
mod <- betadisper(dist_BC, group)
mod

## Perform test
anova(mod)

## Permutation test for F
permutest(mod, pairwise = T)

## Tukey's Honest Significant Differences
(mod.HSD <- TukeyHSD(mod))
plot(mod.HSD)

## Plot the groups and distances to centroids on the
## first two PCoA axes
plot(mod)

## Draw a boxplot of the distances to centroid for each group
boxplot(mod)

df <- data.frame(Distance_to_centroid=mod$distances,Group=mod$group) %>% 
  mutate(Season = case_when(Group == 5 ~ "Spring",
                            Group == 9 ~ "Fall"))
groups <- mod$group


## no difference in homogeneity of dispersion, can continue with the seasonality differences being due to compositional differences
## rather than being driven by differences in composition within sample groups (higher vs lower beta diversity)
```

```{r permdist boxplot, echo = F, warning=F, fig.width=4, fig.height=2.5}
ggplot(data=df,aes(x=Season,y=Distance_to_centroid,fill = Season))+
  geom_boxplot() +
   scale_x_discrete(limits=c("Spring","Fall"))+
  scale_fill_manual(values = c("darkgray", "black"))+
  theme_classic()+
  labs(y = "Distance to centroid")+
  theme(legend.position = "none") 

anova(mod)
```

The PERMANOVA results can be affected when the multivariate dispersions are different across groups. To check this, we calculated multivariate dispersions, which are the distances of observations to their centroids, using the betadisper function of vegan. We compared the mean dispersion between the season groups using the permutest function with 999 permutations. We did not detect significant difference in dispersion, which indicates the spread of the communities (or beta diversity) is similar between seasons.


```{r distance to edge set up, echo = F, include = F}
## GIS dataframe created by Meghan
GIS_select <- GIS_df %>% 
  select(Site_Code, DistanceToEdgeOfPreserve)

data_MH_all <- data_MH_clean %>% 
  #filter(`Collection Method` == "Pit") %>% 
  gather("Species", "abundance", -c(Site:`Collector (1=SES)`,Site_Code, `Inside or Outside (Inside=1)`)) %>% 
  left_join(native_classification, by = "Species") %>% 
  filter(abundance != 0) ## have to drop species that weren't found in the trap, since this is based on sp occurrence
  ## that is pooled at the Replicate level within site codes

data_MH_all_summary <- data_MH_all %>% 
  group_by(Site, `Specific Site Number`, Site_Code, Replicate, `Collection Method`) %>% 
  summarize(total_species = length(Species),
            total_native = sum(Native),
            total_non_native = total_species-total_native,
            prop_non_native = total_non_native/total_species) %>%
  mutate(non_native_binom = ifelse(total_non_native >0, 1, 0)) %>% 
  left_join(GIS_select) %>% 
  mutate(Site_ID = `Specific Site Number`,
         dist_km = (DistanceToEdgeOfPreserve/1000)) %>% ## convert to km since glmer is having trouble with the scale of the data
filter(dist_km != "NA")


## Binomial model of presence of non-native ant sp in all trap types by distance to preserve edge (Random effect of SITE)
## could also do model fitting

glmer_all_0 <- glmer(cbind(total_non_native, total_native) ~ (1|Site_ID), family = "binomial", data = data_MH_all_summary)
glmer_all_fit <- glmer(cbind(total_non_native, total_native) ~ dist_km + (1|Site_ID), family = "binomial", data = data_MH_all_summary)

summary(glmer_all_0)
summary(glmer_all_fit)

## model with the fixed effect of distance + random effect has the lower AIC value and the majority of the weight
AICtab(glmer_all_0, glmer_all_fit, weights = T)
anova(glmer_all_0, glmer_all_fit, test="Chisq")

plot(x=data_MH_all_summary$dist_km, y = data_MH_all_summary$non_native_binom)
MyData =data.frame(cbind(dist_km = seq(from = min(data_MH_all_summary$dist_km), to = max(data_MH_all_summary$dist_km, length.out = nrow(data_MH_all_summary))),
                         Site_ID = seq(from = min(data_MH_all_summary$dist_km), to = max(data_MH_all_summary$dist_km, length.out = nrow(data_MH_all_summary)))))
Pred <- predict(glmer_all_fit, newdata = MyData, type = "response", allow.new.levels=T)
lines(MyData$dist_km, Pred)

## make this in ggplot

pd <-with(data_MH_all_summary,
          data.frame(cbind(dist_km = seq(min(data_MH_all_summary$dist_km), max(data_MH_all_summary$dist_km), length.out = nrow(data_MH_all_summary))),
                     Site_ID = seq(min(data_MH_all_summary$dist_km), max(data_MH_all_summary$dist_km), length.out = nrow(data_MH_all_summary))))

pd1 <- cbind(pd, predict(glmer_all_fit, newdata=pd, type = "response", allow.new.levels = T)) %>% 
  rename(fit_y = `predict(glmer_all_fit, newdata = pd, type = "response", allow.new.levels = T)`)


## binomial model fit to presence/absence of any non-native ant sp found in pitfall trap
fig_binom<- ggplot(data_MH_all_summary, aes((dist_km), non_native_binom), shape = `Collection Method`)+
  geom_point(aes(shape = `Collection Method`),position = position_jitter(width =.02, height = 0.02), alpha = .6)+
  geom_line(data = pd1, aes(x = dist_km, y = fit_y))+
  theme_classic()+
  labs(x = "Distance to edge of preserve (km)",
       y = "Probability of non-native ant species present")

## comfortable with the spread of the different collection methods across the distance to edge,
## and occurrence of non-native ant spp
ggplot(data_MH_all_summary, aes((dist_km), non_native_binom))+
  geom_point(position = position_jitter(width =.02, height = 0.02), alpha = .6)+
  geom_line(data = pd1, aes(x = dist_km, y = fit_y))+
  theme_classic()+
  labs(x = "Distance to edge of preserve (km)",
       y = "Probability of non-native ant species present")



##### visualize and test for relationship between the to dominant non-native ants and distance to edge
data_MH_all_invasives <- data_MH_clean %>% 
  #filter(`Collection Method` == "Pit") %>% All collection methods
  gather("Species", "abundance", -c(Site:`Collector (1=SES)`,Site_Code, `Inside or Outside (Inside=1)`)) %>% 
  left_join(native_classification, by = "Species") %>% 
  filter(#abundance != 0, ## keeping in samples for which we did not observe the two spp of interest
         Species == "Nylanderia fulva" |
           Species == "Solenopsis invicta") %>% 
  left_join(GIS_select) %>% 
  filter(DistanceToEdgeOfPreserve != "NA") %>% 
  mutate(Site_ID = `Specific Site Number`,
         dist_km = (DistanceToEdgeOfPreserve/1000))

lmer_inv_0 <- lmer(abundance ~ (1|Site_ID), data = data_MH_all_invasives, REML=F)
lmer_inv_fit <- lmer(abundance ~ dist_km + (1|Site_ID), data = data_MH_all_invasives, REML=F)

summary(lmer_inv_0)
summary(lmer_inv_fit)

library(MuMIn)

r.squaredGLMM(lmer_inv_fit)

## model with the fixed effect of distance has the lower AIC value and the majority of the weight
AICtab(lmer_inv_0, lmer_inv_fit, weights = T)
dist_edge_lmer <- anova(lmer_inv_0, lmer_inv_fit, test="Chisq")

pd_inv <-with(data_MH_all_invasives,
          data.frame(cbind(dist_km = seq(min(data_MH_all_invasives$dist_km), max(data_MH_all_invasives$dist_km), length.out = nrow(data_MH_all_invasives))),
                     Site_ID = seq(min(data_MH_all_invasives$dist_km), max(data_MH_all_invasives$dist_km), length.out = nrow(data_MH_all_invasives))))

pd1_inv <- cbind(pd_inv, predict(lmer_inv_fit, newdata=pd_inv, type = "response", allow.new.levels = T)) %>% 
  rename(fit_y = `predict(lmer_inv_fit, newdata = pd_inv, type = "response", allow.new.levels = T)`)



## similar trend across collection methods -- OK to "pool" 
ggplot(data_MH_all_invasives,aes(dist_km, abundance))+
  geom_point(aes(color = Species, shape = `Collection Method`))+
  geom_line(data = pd1_inv, aes(x = dist_km, y = fit_y))+
  #geom_smooth(method = "lm", se=F) ## nearly identical fit to our estimate with random effects of site
  theme_classic()

dist_edge_fig_non_natives <-ggplot(data_MH_all_invasives,aes(dist_km, abundance))+
  geom_jitter(aes(color = Species, shape = `Collection Method`), width = 0.05, height = 1, alpha = .75, size =2.5)+
    geom_line(data = pd1_inv, aes(x = dist_km, y = fit_y))+
  scale_color_manual(values = c("tan", "steelblue4"))+
  theme_classic()+
  #theme(legend.position = "bottom")+
  labs(x = "Distance to preserve boundary (km)",
       y = "Ant abundance")
  #geom_smooth(method = "lm", se=F)

## combining across collection methods doesn't affect presence/absence probability of presence,
## but it does greatly increases the abundances

##### visualize and test for relationship between the to dominant non-native ants and distance to edge
data_MH_all_invasives_no_zero <- data_MH_clean %>% 
  #filter(`Collection Method` == "Pit") %>% All collection methods
  gather("Species", "abundance", -c(Site:`Collector (1=SES)`,Site_Code, `Inside or Outside (Inside=1)`)) %>% 
  left_join(native_classification, by = "Species") %>% 
  filter(abundance != 0, ## drop instances where we did not observe species
    Species == "Nylanderia fulva" |
      Species == "Solenopsis invicta") %>% 
  left_join(GIS_select) %>% 
  filter(DistanceToEdgeOfPreserve != "NA") %>% 
  mutate(Site_ID = `Specific Site Number`,
    dist_km = (DistanceToEdgeOfPreserve/1000)) ## convert to km since glmer is having trouble with the scale of the data

lmer_inv_0 <- lm(abundance ~ (1), data = data_MH_all_invasives_no_zero)
lmer_inv_fit <- lm(abundance ~ dist_km, data = data_MH_all_invasives_no_zero)
##^^ getting an error message "boundary (singular)" means variance is basically = 0 -- not fitting well
summary(lmer_inv_0)
summary(lmer_inv_fit)

## model with the random effect has the lower AIC value and the majority of the weight -- go with model that includes zeros
AICtab(lmer_inv_0, lmer_inv_fit, weights = T)
anova(lmer_inv_0, lmer_inv_fit)

plot(x=data_MH_all_invasives_no_zero$dist_km, y = data_MH_all_invasives_no_zero$abundance)
MyData =data.frame(cbind(dist_km = seq(from = min(data_MH_all_invasives_no_zero$dist_km), to = max(data_MH_all_invasives_no_zero$dist_km, length.out = nrow(data_MH_all_summary))),
                         Site_ID = seq(from = min(data_MH_all_invasives_no_zero$dist_km), to = max(data_MH_all_invasives_no_zero$dist_km, length.out = nrow(data_MH_all_summary)))))
Pred <- predict(lmer_inv_fit, newdata = MyData, type = "response", allow.new.levels=T)
lines(MyData$dist_km, Pred)

## make this in ggplot

pd <-with(data_MH_all_invasives_no_zero,
          data.frame(cbind(dist_km = seq(min(data_MH_all_invasives_no_zero$dist_km), max(data_MH_all_invasives_no_zero$dist_km), length.out = nrow(data_MH_all_invasives_no_zero))),
                     Site_ID = seq(min(data_MH_all_invasives_no_zero$dist_km), max(data_MH_all_invasives_no_zero$dist_km), length.out = nrow(data_MH_all_invasives_no_zero))))

pd1 <- cbind(pd, predict(lmer_inv_fit, newdata=pd, type = "response", allow.new.levels = T)) %>% 
  rename(fit_y = `predict(lmer_inv_fit, newdata = pd, type = "response", allow.new.levels = T)`)


## similar trend across collection methods -- OK to "pool" 
ggplot(data_MH_all_invasives_no_zero,aes(dist_km, abundance))+
  geom_point(aes(color = Species, shape = `Collection Method`))+
  theme_classic()+
  geom_smooth(method = "lm", se=F)+
  facet_grid(~`Collection Method`)

ggplot(data_MH_all_invasives_no_zero,aes(dist_km, abundance))+
  geom_point(aes(color = Species))+
  theme_classic()+
  geom_smooth(method = "lm", se=F)
```

## Probability of non-native ants decreases within the preserve
```{r dist to edge binom fig, echo = F, warning = F, fig.height=4, fig.width = 4}
fig_binom+
  theme(legend.position = "bottom")
anova_glmer_binom <- anova(glmer_all_0, glmer_all_fit, test="Chisq")
```

```{r table, echo = F, warning = F}
kable(anova_glmer_binom)
```


Each point represents whether or not a non-native ant species was detected in the sample, shapes represent the method of collection (Hand = circles, Pitfall traps = triangles, and Winkler = squares). The fitted line is the Binomial model estimates for the probability of the presence of a non-native ant species within the trap. This was an improved fit with site as a random effect and distance to preserve edge compared to that of the intercept with the random effect of site.  

## Abundances of *S. invicta* and *N fulva* at each collection site

```{r dist to edge fig, echo = F, warning = F, fig.height =2.5, fig.width = 5}
dist_edge_fig_non_natives
```

```{r dist to edge table, echo = F, warning = F}
kable(dist_edge_lmer)
```


Here, each point represents the abundance of *N. fulva* (tan) or *S. invicta* (blue) by collection method (Hand = circles, Pitfall traps = triangles, Winkler = squares) at a given sampling site that is some distance (km) to the nearest preserve boundary. The line is the best fit regression ($R^2$ = .026) with site as a random effect and distance to edge as a fixed effect. However, this was only a slight improvement over the null model (intercept and site as a random effect). (*To me, this suggests that distance to edge of preserve is not a strong predictor of abundances of these two ant species and that this is most likley due to the high number of zeros and very low abundances in the data set. I think we could cut this figure and just include the stronger result (above) from the binomial model.*)

## Next steps
- Decide whether to classify *P. falvens* complex as native or non-native. -- IT SHOULD BE NON-Native
- Decide whether it makes more sense to present alpha diversity indices based on *occurrence* or *abundance*. (*I think abundance makes more sense, since the Shannon and Simpson indices are abundance based*).

